---
title: "HRM LakeWatchers: Trends in Urban Lake Water Quality"
subtitle: "Halifax Regional Municipality"
author: "Vinura Jayowin Amarasinghe"
date: today
---

## Introduction

Halifax Regional Municipality manages an extensive network of urban lakes
used for recreation, drinking water supply, and ecosystem services. Recent
summers have seen repeated blue-green algae advisories at supervised
beaches, raising questions about long-term trends in nutrient enrichment
and warming.

This analysis uses the open
[HRM LakeWatchers dataset](https://doi.org/10.25976/85pw-zj92)
on DataStream to examine multi-year trends in key water quality
indicators.

> **Note:** This report is currently a scaffold. The data fetch script is
> in place but the analysis sections will be populated once the
> DataStream API key arrives and the initial data exploration is complete.

## Data

```{r}
#| label: setup
#| message: false
#| eval: false

library(tidyverse)
library(lubridate)
library(Kendall)

# Will load processed data once R/02_clean_data.R has been run
# lakes <- read_csv("data/processed/lakewatchers_clean.csv")
```

## Site overview

*Map of monitoring locations and summary of sites — to be added.*

## Temperature trends

*Time-series plots per lake, Mann-Kendall test, brief interpretation —
to be added.*

## Dissolved oxygen patterns

*Box plots by lake and season; flag exceedances of relevant guidelines —
to be added.*

## Nutrient enrichment

*Total phosphorus and chlorophyll-a trends; trophic-status classification —
to be added.*

## Discussion

*Short — what the data shows, what the limitations are, what would be
worth investigating further. To be added.*

## Methods notes

*Mann-Kendall justification, handling of detection limits, missing data
treatment — to be added.*

## Data citation

> Halifax Regional Municipality (2025). HRM LakeWatchers (dataset). 2.0.0.
> DataStream. https://doi.org/10.25976/85pw-zj92

## Session info

```{r}
#| code-fold: true
sessionInfo()
```